---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
```

While we could use other methods such as traitar and/or PHENDB for metagneomic data, 
16S data requires a reference-based approach. For 16S data, one could likely extrapolate from 
pfam or eggnog annotations obtained with picrust2. Here we use BacDive--a 
database of bacterial phenotypic information--to obtain phenotype data derived from
experimental work.

Install BacDive API access package
```{r}
#install.packages("BacDive", repos="http://R-Forge.R-project.org")
```

Load BacDive
```{r}
library(BacDive)
```

Get BacDive auth.
```{r}
bacdive <- BacDive::open_bacdive("kiledal@umich.edu",password = "sGcBn2$P&iTz#Y")
```

Function for getting oxygen tolerance
```{r}
cols <- c("ra_and_subs","reads_and_subs","reads","rank","NCBI_id","tax","full_lineage","lineage")

mmseqs <- read_tsv("data/mmseqs/concreteMetaG_report_w_standardized_lineage", 
                   col_names = cols,trim_ws = F)

tax_levels <- c("root", "kingdom", "phylum", "class", "order", "family","genus","species")

tax_profile <- mmseqs %>% 
  mutate(rel_abund = reads / sum(reads)) %>% 
  filter(!tax %in% c("unclassified","  cellular organisms")) %>% 
  mutate(lineage = str_glue("r__Root;{lineage}"),
         lineage = str_remove_all(lineage, "(;[a-z]__)*$")) %>% 
  filter(!str_detect(lineage, ";NA")) %>% 
  separate(lineage, into = tax_levels, remove = FALSE,sep = ";") %>% 
  mutate(across(tax_levels, ~str_remove(.x, "[a-z]__")))

genus_abund <- tax_profile %>% 
  group_by(kingdom,phylum,class,order, family, genus) %>% 
  summarise(rel_abund = sum(rel_abund)) %>% 
  arrange(desc(rel_abund))

unique_genera <- genus_abund %>% 
  filter(!is.na(genus) & genus != "") %>% 
  pull(genus) %>% unique()
  

genus <- "Campylobacter"
genus <- "this_is_a_fake_name"
i = 30


get_O2_tolerance <- function(genus){
  
  #ids <- request(bacdive, genus , "taxon")
  
  data <- retrieve(object = bacdive, query = genus,search = "taxon", sleep = 0.1)
  
  if (length(data) > 0) {
  
    has_o2_tolerance <- numeric()
    for (i in (1:length(data))) {
      if (!is.null(data[[i]][["Physiology and metabolism"]][["oxygen tolerance"]])) {
        has_o2_tolerance <- c(has_o2_tolerance,i)
      }
    }
    
    data2 <- data[has_o2_tolerance]
    
    if (length(data2) > 0) {
    
      extracted_oxygen_tolerance_ids <- purrr::map(data2,~.x$`Physiology and metabolism`$`oxygen tolerance`$`@ref`) %>% 
        unlist() %>% 
        as.character() %>% 
        data.frame(id = .)
      
      extracted_oxygen_tolerance <- purrr::map(data2,~.x$`Physiology and metabolism`$`oxygen tolerance`$`oxygen tolerance`) %>% 
        unlist() %>% 
        as.character() %>% 
        data.frame(oxygen_tolerance = .) #%>% 
        bind_cols(extracted_oxygen_tolerance_ids) %>% 
        mutate(genus = genus)
    } else return(data.frame(genus = genus, oxygen_tolerance = "not_found"))
    
    return(extracted_oxygen_tolerance)
  } else return(data.frame(genus = genus, oxygen_tolerance = "not_found"))
  
}
```



```{r}
example_oxy_tolerance <- map_df(unique_genera[1:1000],get_O2_tolerance)

saveRDS(example_oxy_tolerance, "data/bacdive_oxy_tolerance.rds")

example_oxy_tolerance <- read_rds("data/bacdive_oxy_tolerance.rds")

oxy_tolerance_wide <- example_oxy_tolerance %>% 
  group_by(genus, oxygen_tolerance) %>% 
  summarise(count = n()) %>% 
  pivot_wider(names_from = oxygen_tolerance, values_from = count) %>% 
  right_join(genus_abund %>% select(any_of(tax_levels), rel_abund)) %>% 
  relocate(any_of(tax_levels), rel_abund)

top_oxy_tolerance_wide <- example_oxy_tolerance %>% 
  group_by(genus, oxygen_tolerance) %>% 
  summarise(count = n()) %>% 
  group_by(genus) %>% 
  slice_max(count, n = 1,with_ties = FALSE) %>% 
  pivot_wider(names_from = oxygen_tolerance, values_from = count) %>% 
  left_join(genus_abund %>% select(any_of(tax_levels), rel_abund))    


rel_abund_w_ox_tolerance <- example_oxy_tolerance %>% 
  group_by(genus, oxygen_tolerance) %>% 
  summarise(count = n()) %>% 
  group_by(genus) %>% 
  slice_max(count, n = 1,with_ties = FALSE) %>% 
  left_join(genus_abund %>% select(any_of(tax_levels), rel_abund))

rel_abund_by_ox_tolerance <- rel_abund_w_ox_tolerance %>% 
  group_by(oxygen_tolerance) %>% 
  summarise(abund = sum(rel_abund) %>% round(5))

rel_abund_by_ox_tolerance %>% 
  filter(oxygen_tolerance != "not_found") %>% 
  pull("abund") %>% 
  sum()

```


```{r}
rel_abund_by_ox_tolerance %>% 
  mutate(other = 1 - sum(abund),
         abund = if_else(oxygen_tolerance == "not_found", abund + other, abund)) %>% 
  ggplot(aes(1, abund, fill = oxygen_tolerance)) +
  geom_bar(stat = "identity")

rel_abund_by_ox_tolerance %>% 
  mutate(other = 1 - sum(abund),
         abund = if_else(oxygen_tolerance == "not_found", abund + other, abund)) %>% 
  filter(oxygen_tolerance != "not_found") %>% 
  ggplot(aes(1, abund, fill = oxygen_tolerance)) +
  geom_bar(stat = "identity")


```









