---
title: "kegg_enrichment"
author: "Anders Kiledal"
date: "11/12/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
library(tidyverse)
```


```{r}
s3_kegg_module <- read_tsv("data/shogun/shogun_burst_function/.strain.kegg.modules.txt")

s3_kegg_module_coverage <- read_tsv("data/shogun/shogun_burst_function/.strain.kegg.modules.coverage.txt")

kegg_modules <- read_tsv("data/reference/shogun/function/ko-module-annotations.txt")

```

```{r}
s3_kegg_pathway <- read_tsv("data/shogun/shogun_burst_function/.strain.kegg.pathways.txt")

s3_kegg_pathway_coverage <- read_tsv("data/shogun/shogun_burst_function/.strain.kegg.pathways.coverage.txt")

kegg_pathways <- read_tsv("data/reference/shogun/function/ko-pathway-annotations.txt")


kegg_enzymes <- read_tsv("data/reference/shogun/function/ko-enzyme-annotations.txt")

kegg_strain <- read_tsv("data/reference/shogun/function/ko-strain2ko.txt")

ko_info <- read_delim("data/reference/ko_info.tsv", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE) %>% 
  rename(ko = "X1",
         ko_name = "X2")

```



The way Mengyin calculated odds ratios was:

Odds ratios were calculated for each COG category identified in
the metagenomic data set by calculating the ratio (A/B)/(C/D), where A is
the number of reads in the metagenome that are in a given COG category,
B is the number of reads in the metagenome that are in all of the other
COG categories, C is the number of proteins in the COG database that are
in a given COG category, and D is the number of proteins in the COG
database that are in all of the other COG categories. COG categories with
odds ratios of 1 were considered enriched (42), meaning that they are
more highly represented in the metagenomic data than in the protein
database.

Parse the kegg data
```{r}

#res <- readLines("data/reference/")
#R> length(res)

ko00001 <- read_delim("data/reference/ko00001.keg", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE) %>% 
  separate(X1, into = c("level","id","name"),sep = "      ") %>% 
  mutate(id = if_else(str_detect(level,"^A"),str_remove(level,"^A"),id),
         level = if_else(str_detect(level,"^A"),"A",level),
         id = if_else(str_detect(level,"^B"),str_remove(level,"^B  "),id),
         level = if_else(str_detect(level,"^B"),"B",level),
         id = if_else(str_detect(level,"^C"),str_remove(level,"^C    "),id),
         level = if_else(str_detect(level,"^C"),"C",level)
         ) %>% 
  filter(id != "B") %>% 
  separate(id,into = c("id","name"),sep = "  ") %>% 
  mutate(name = if_else(level %in% c("A","B","C"),str_remove(id,"[0-9]* "),name),
         id = if_else(level %in% c("A","B","C"),str_extract(id,"[0-9]*"),id),
         A = "",
         B = "",
         C = "",
         family = "",
         path_type = "",
         pathway = "")


A=""
B=""
C=""


ko_process <- function(table) {
  
  in_table <- table
  for (i in 1:nrow(in_table)){
    if (in_table[i,1]=="A"){
      A <- in_table[i,"id"]
      family <- in_table[i,"name"] 
      
      #Reset lower values
      B <- ""
      c <- ""
      path_type = ""
      pathway = ""
      
    }
    if (in_table[i,1]=="B"){
      B <- in_table[i,"id"]
      path_type <- in_table[i,"name"]
      
      C <- ""
      pathway <- ""
    }
    if (in_table[i,1]=="C"){
      C <- in_table[i,"id"]
      pathway <- in_table[i,"name"]
    }
    
    in_table[i,"A"] <- A
    in_table[i,"B"] <- B
    in_table[i,"C"] <- C
    in_table[i,"family"] <- family
    in_table[i,"path_type"] <- path_type
    in_table[i,"pathway"] <- pathway
  }
  return(in_table)
}

kegg_data <- ko_process(ko00001)



```


```{r}

kegg <- kegg_data %>% 
  filter(level == "D") %>% 
  mutate(total = n()) %>% 
  group_by(pathway) %>% 
  mutate(n_in = n(),
         n_out = total - n_in)


meta_kegg <- s3_kegg_pathway <- read_tsv("data/shogun/burst/taxatable.strain.kegg.txt") %>% 
  rename(id = "#KEGG ID") %>% 
  mutate(meta_abund = S3_Fallen_R1 + S3_Fallen_R1) %>% 
  select(id, meta_abund)


joined_kegg <- kegg %>% 
  left_join(meta_kegg) %>% 
  ungroup() %>% 
  mutate(meta_abund = replace_na(meta_abund,0)) %>% 
  mutate(meta_total = sum(meta_abund)) %>% 
  group_by(pathway) %>% 
  mutate(meta_in = sum(meta_abund),
         meta_out = meta_total - meta_in,
         OR = (meta_in/meta_out) / (n_in/n_out))


pathway_or <- joined_kegg %>% 
  select(pathway, OR) %>% 
  distinct()


#Kegg pathway abundance
s3_kegg_pathway <- read_tsv("data/shogun/burst/taxatable.strain.kegg.pathways.txt")

```






```{r}
s3_mod <- read_tsv("data/shogun/burst/taxatable.strain.kegg.modules.txt")

mod_info <- read_tsv("data/reference/KEGG_modules_info.tsv",col_names = c("module","name")) %>% distinct()

get_module_ko <- function(module) {
  data <- tryCatch(KEGGREST::keggGet(module), error=function(e) NULL)
  res <- str_replace_all(data[[1]]$DEFINITION,"[\\+(),-]"," ") %>% 
    str_split("\\s+") %>% 
    unlist() %>% 
    Filter(function(x){str_detect(x, "^K")}, .) 
  
  if(length(res) > 0){
  results <- res %>% 
    data.frame(mod = module,
               ko = .)
  }
  if(length(res) == 0){
  results <- data.frame(module = module,
               ko = NA)
  }
  
  return(results)
  }


mod_test_list <- c("M00014","M00095")

results <- lapply(unique(mod_info$module), get_module_ko)

re2 <- bind_rows(results) %>% rename(module = "mod")

write_tsv(re2, "data/reference/kegg_module_mem.tsv")



mod_W_abund <- re2 %>% 
  left_join(meta_kegg %>% rename(ko = "id")) %>% 
  left_join(mod_info %>% rename(module_name = "name")) %>% 
  left_join(ko_info) %>% 
  mutate(meta_abund = replace_na(meta_abund,0)) %>% 
  mutate(meta_total = sum(meta_abund)) %>% 
  mutate(total = n()) %>% 
  group_by(module) %>% 
  mutate(n_in = n(),
         n_out = total - n_in) %>% 
  filter(meta_abund > 0) %>% 
  group_by(module) %>% 
  mutate(meta_in = sum(meta_abund),
         meta_out = meta_total - meta_in,
         OR = (meta_in/meta_out) / (n_in/n_out))


pathway_or <- joined_kegg %>% 
  select(pathway, OR) %>% 
  distinct()



mod_test <- keggGet("M00014")

test_vec <- str_replace_all(mod_test[[1]]$DEFINITION,"[(),-]"," ") %>% str_split("\\s+") %>% unlist()

filt_fun <- function(x){str_detect(x, "^K")}

dat <- get_module_ko("M00095")
```


```{r}
mod_test <- keggGet("M00014")
```


